---
title: "PA1_template.Rmd"
author: "Sandra Cuffee"
date: "November 15, 2015"
output: html_document
---

 This assignment makes use of data from a personal activity monitoring device.
 This device collects data at 5 minute intervals through out the day.
 The data consists of two months of data from an anonymous individual collected during the
 months of October and November, 2012 and include the number of stepstaken in 5 minute
 intervals each day.
 The variables included in this dataset are:
 steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)
 date: The date on which the measurement was taken in YYYY-MM-DD format
 interval: Identifier for the 5-minute interval in which measurement was taken
 The dataset is stored in a comma-separated-value (CSV) file and
 there are a total of 17,568 observations in this dataset.


Add packages to execute script

```{r}

library(data.table)
library(plyr)
library(dplyr)
library (tidyr)
library(stringr)
library(graphics)
library(knitr)
```

 Load the function
 
```{r data_analysis, echo = TRUE}

  data_analysis <- function()
  
  {     
      #Check to see if requred directory exists, if not create it.
    
      if(!file.exists("C:/Users/Owner/Documents/specdata"))
         {
              dir.create("C:/Users/Owner/Documents/specdata")
         } 
  
  
      # download te data
    
      fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
      download.file(fileUrl, destfile = "C:/Users/Owner/Documents/specdata/repdata_data_activity.zip", method = "libcurl")
      unzip("C:/Users/Owner/Documents/specdata/repdata_data_activity.zip")
      
      #Process the data with the NA's removed
     
      activityData <- na.omit(read.csv("C:/Users/Owner/Documents/specdata/activity.csv")) 
      
      # Calculate the daily steps
      activityDataSteps <- group_by(activityData, date)
      activityDataSteps 
      
      # Calulate Create histogram representing the daily number of steps
      
    
      hist(activityDataSteps$steps,main = " Daily Steps", xlab = "Steps", col ="blue")
      
      
      # Caluculate the mean number of daily steps
      
      activityDataSteps1 <- summarize(activityDataSteps, Mean = mean(steps))
      activityDataSteps1  
      meanSteps <- mean(activityData$steps)
      print ("The mean steps are:")
      print (meanSteps)
      
  
      # Create histogram representing the daily mean number of daily steps
      
      
      hist(activityDataSteps1$Mean, main = "Mean Daily Steps", xlab = "Steps", col ="blue")
      
      
      #Caluculate the median daily steps
      
      activityDataSteps2 <- summarize(activityDataSteps, Median = median(steps))
      activityDataSteps2        
      medianSteps <- (median(activityData$steps))
      print ("The median steps are:")
      print(medianSteps)
      
      # Create histogram representing the daily median number of daily steps
      
      
      hist(activityDataSteps2$Median, main = "Median Daily Steps", xlab = "Steps", col ="blue")
      
      #Caluculate the interval steps
      
       activityDataSteps5 <- group_by(activityData, as.factor(interval))
      
      #A time series plot 5-minute interval (x-axis), average number of steps taken, averaged across all days (y-axis)
     
     plot(as.numeric(activityDataSteps5$interval),activityDataSteps5$steps, type = "l", xaxt = "n", main = "Daily Steps by Intervals", xlab = "Time Intervals", ylab = "Number of Steps by Interval", col = "blue")
      axis(1, at=as.numeric(activityDataSteps5$interval), labels = activityDataSteps5$interval)
      
     
     
      
   
      #Calculate and report the total number of missing values in the dataset
      
      activityData1 <- read.csv("C:/Users/Owner/Documents/specdata/activity.csv")
      NAS <- sum(is.na(activityData1$steps))
      print("The number of NA's are: ")
      print(NAS)
      
       #Imputing missing values      
      
      
      #sStrategy for filling in all of the missing values in the dataset will be to use the mean value
      activityDataSteps6 <- replace(activityData1$steps, is.na(activityData1$steps), mean(activityData1$steps, na.rm = TRUE)) 
    
      
      
      
      
      
  }    
  
```


```{r Execute the function, figure}

    data_analysis()
    
```
  